\chapter*{Acknowledgements}

First of all, I would like to thank my supervisor PhD Robert Dąbrowski for expert guidance during the "Software Optimization, Visualization and Analysis" (SOVA) seminars. His help in the university administrative procedures involving this thesis, overall assistance and patience in overcoming various obstacles by me was invaluable.

Another lecturer present during the seminars was Prof. Krzysztof Stencel who provided me with essential advice. I greatly appreciate his willingness to share critical insight related to my thesis and the whole defence process.

Apart from the leaders of the seminars, I must mention my colleagues who were attending my presentations. This was the time when we shared our remarks on each other's work. My thesis would not be the same without their input during these discussions.

From my colleagues, I must single out and thank Małgorzata E. Z. Galińska who was the first to test my VS Code plugin and assisted me with debugging it.

I thank my former employers and dear friends Mathias Åsberg (CEO \href{https://newnative.ai/}{NewNative}) and Radosław Dembkowski (CEO \href{https://www.radcode.co/}{RadCode}) for their efforts at helping me gather the data.

I applaud everyone who decided to share their data to proceed with my research project. It would be a cliché to say that my thesis would not exist without them.

I am eternally grateful for the Faculty of Mathematics, Mechanics and Informatics of the University of Warsaw for giving me the chance to study Computer Science and conduct my research there. Every year spent at the Faculty was an inimitable adventure for me.

Last but not least, my thanks go to my friends for their indispensable presence and my family for their immense patience and unwavering support both emotional, and financial.

\begin{flushright}
Warsaw, December 2022

Mateusz Bajorek
\end{flushright}
